Score-based tests of measurement invariance: Use in practice

### 2013-09-26 frontiers/frontiers.Rnw
make changes according to Ed's comments.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@267 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\readme{We need a better intro here.  The special topic is about measurement invariance, so most of this paragraph is obvious to everyone.  Maybe start with a remark about approximate invariance and try to tie our topic to that somehow.}
Parameter invariance is a fundamental assumption for measurement model in 
developing psychometric scales. Violation of it could happen in any estimated parameter.  
In order to fix the psychometric scale, we need to specify the violating parameter.  
Moreover, in practice model misspecification is an inevitable issue to a certain extent 
and may interfere with parameter invariance specificity. A family of tests derived from 
stochastic process has been recently proposed to identify specific parameters impacted 
by measurement invariance violation. However, these tests have only been used to a limited 
type of parameter violation. It is this paper's intent to provide a detailed examination 
of the tests' performance under practical research scenarios.

@@

## remove cases with imputed values(not in 1,...,9)
compcases <- apply(YouthGratitude[,4:28], 1, function(x) all(x %in% 1:9))
yg <- YouthGratitude[compcases,]
@
Next, we fit two models in lavaan: a one-factor model where loadings are restricted 
to be equal across age groups, and a one-factor model where loadings are free across age groups.
<<modest>>=
## Restricted model
reduced <- cfa("f1=~gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
             data=yg, group="agegroup", meanstructure=TRUE,
             group.equal="loadings")

## Unrestricted model
full <- cfa("f1=~gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
             data=yg, group="agegroup", meanstructure=TRUE)
@
Finally, we can get the results of a likelihood ratio test via the \verb+anova()+ 
function, which implies that the GQ-6 violates measurement invariance.
<<LRT>>=
## LRT
anova(full, reduced)
@

\readme{I just added LMuo to the sctest command and corrected the command below; the range statistic is different.}
Another similar statistics is $\mathit{LM}_\mathit{uo}$ as shown in Equation~ \eqref{eq:lmuo}. We can obtain this statistic by using the following function. 
<<LM_UO>>=
## unordered LM
lm_uo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="LMuo")
round(lm_uo$p.value, 4)
@
The above command specifies that we wish to test model parameter numbers 1 through 4, which are the factor loadings supplied by lavaan.  The ordering of parameters can be obtained via the command \verb+coef(reduced)+.

Because our sample size is large, this likelihood ratio test is known to be sensitive to 
small measurement invariance violations \cite{Ben80}.  That is, the above tests 
are sensitive to small measurement invariance violations that are not likely to be 
of interest to researchers.  For example, imagine that the 15-year-olds' 
parameters are slightly different than the other age groups.  
The 15-year-olds are in the middle of the age groups, and there 
is not likely to be any theoretical justification for 15-year-olds 
differing from every other age group.  One solution to this problem would be the {\em approximate invariance} methods proposed by Muth\'{e}n and others.\readme{Ting: Can you get some approximate invariance references and put them here?}
Alternatively, we can use the proposed family of score-based statistics to obtain tests that are sensitive to the ordering of age.

@@

<<Get Estimate>>=
dm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="DM")
cvm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="CvM")
maxlm <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLM")
round(c(dm$p.value, cvm$p.value, maxlm$p.value), 2)
@
%% move this code description to the cat2BB part. 
%In the code above, the \verb+parm=1:4+ argument means that 
%we wish to test model parameter numbers 1 through 4 (which are the factor loadings), 
%which are supplied by lavaan.  The ordering of parameters can be obtained via the 
%command \verb+coef(reduced)+.  
We see that two of the three $p$-values output at 
the end of the code are larger than that associated with the LRT (with the CvM statistic being non-significant).

@@

To carry out the tests, we can rely on the same function that we used
for the continuous test statistics.  As mentioned previously,
calculation of the ``maxLMo'' statistic (Equation~\eqref{eq:maxlmo})
can be lengthy from the need to simulate critical values.
%% This is more complicated than necessary, but it is written like
%% this so that we can load the ordinal critical values behind the
%% scenes without confusing the reader.
<<critvals,echo=FALSE,strip.white=all>>=
if(file.exists("opval1.rda")){
  load("opval1.rda")
} else {
  gefp1 <- gefp(reduced, fit=NULL, order.by=as.numeric(yg$agegroup),
                vcov=info_full, sandwich=FALSE, parm=1:4)
  opval1 <- ordL2BB(gefp1)
  save(opval1, file="opval1.rda")
}
## Get test statistics using fnl=opval1 without showing it in the paper
wdmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="WDMo")
maxlmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLMo",
                 fnl=opval1)
@ 
<<ordtest,eval=FALSE,strip.white=all>>=
wdmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="WDMo")
maxlmo <- sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLMo")
@ 
<<pval,strip.white=all>>=
round(c(wdmo$p.value, maxlmo$p.value), 2)
@ 
In computing the ordinal test statistics, we
obtain~$p=\Sexpr{round(wdmo$p.value,2)}$
and~$p=\Sexpr{round(maxlmo$p.value,2)}$, respectively.
Both $p$-values are clearly larger than that of the 
likelihood ratio test and neither is significant at $\alpha=0.05$. 
This provides evidence that there is no measurement
invariance violation that is monotonic with age group.  Instead, given
the large sample size, the likelihood ratio test may be overly
sensitive to anomalous, non-monotonic violations at one (or a few) age groups.

@@

%Or alternatively, we could use the sctest.lavaan function, but assign ordinal statistics.
\begin{figure}
\caption{Fluctuation processes for the $\mathit{WDM}_o$ statistic (left panel) and the $\max\mathit{LM}_o$ statistic (right panel).}
\label{fig:ordres}
\setkeys{Gin}{width=\textwidth}
<<ordres, fig=TRUE, height=3, width=6, echo=FALSE, results=hide>>=
par(mfrow = c(1, 2), cex = 0.65)
sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="WDMo", 
       fplot=TRUE, ylim=c(0,3.5), xlab="Age group",
       ylab = "Weighted max statistics", axes=FALSE)
axis(1, at = 1:5, labels = c("10-11", "12-13", "14", "15", "16"))
box()

sctest(reduced, as.numeric(yg$agegroup), parm=1:4, stat="maxLMo", 
       fnl=opval1, fplot=TRUE, ylim=c(0,14), xlab="Age group",
       ylab = "LM statistics", axes = FALSE)
axis(1, at = 1:5, labels = c("10-11", "12-13", "14", "15", "16"))
box()
@
\end{figure}

@@

To examine these specificity issues, we generated data from a
two-factor model with three indicators each (see Figure~\ref{fig:path}).
\readme{Please make sure the text below is correct.  The previous text stated that the violation occurs in all loadings and all unique variances.}
The measurement invariance
violation occurred in one of three places: the factor loading associated with Scale 1 
 ($\lambda_{11}$), the unique variance associated with Scale 1
 ($\epsilon_{11}$), or the factor covariance $\sigma_{12}$.  
We then tested for measurement invariance in seven subsets of parameters: the three
individual parameters noted above, all six factor loadings, % $\lambda_{11}--\lambda_{62}$, and 
and all six unique variances. % $\epsilon_{11}--\epsilon_{66}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Power and Type I error were examined across three sample sizes
($n=120,480,960$), three numbers of categories ($m=4,8,12$) and
17 magnitudes of invariance violations. 
The measurement invariance violations began at level $1+m/2$ of $V$
and were consistent thereafter. These
violations were based on the auxiliary variable $V$: individuals below
the $1+m/2$ percentile of $V$ deviated from individuals above the $1+m/2$ 
percentile  by $d$ times the parameters' asymptotic standard errors
(scaled by $\sqrt n$), with $d=0,0,25,0.5,...,4$. 
For each combination of sample size ($n$) $\times$ violation
magnitude ($d$) $\times$ violating parameter $\times$ categories $m$, 5000 datasets were
generated and tested. Four statistics were examined above. \readme{What four statistics?  Equation numbers would be helpful.} 
In all conditions, we
maintained equal sample sizes at each level of the ordinal variable.

\subsection{Results}
\readme{When referring to statistics, it would be good to use the notation from Equations 6--8.  For example, instead of ``unordered $\max\mathit{LM}$'', you would have $\max\mathit{LM}_{uo}$.}
Full simulation results are presented in Figure~\ref{fig:sim11res} to 
Figure~\ref{fig:sim13res}; 
Figure~\ref{fig:sim11res} displays power curves as a function of 
violation magnitude for the first
factor loading $\lambda_{11}$, with panels reflecting the parameters 
being tested and lines reflecting $n$.
Figures~\ref{fig:sim12res} and Figure~\ref{fig:sim13res} display 
similar power curves when the factor covariance 
$\phi_{12}$ and error variance $\epsilon_{11}$ violate measurement
invariance, respectively. 
All three test statistics exhibited 
similar results, with unordered  $\max\mathit{LM}$ demonstrating lower power across all situations. ordered $\max\mathit{LM}$ and double-max exhibited equivalent power curves because, when only one parameter is tested, the statistics are equivalent. \readme{Should say that the unordered statistic is asymptotically equivalent to the LRT.}

@@

In summary, we found that the proposed tests can attribute
measurement invariance violations to the correct parameter.  This
provides evidence that, in practice, one can have confidence in the
tests' abilities to locate the measurement invariance violation.  Of
course, this statement is qualified by the fact that, in this
simulation, the model was correctly specified.  In the following
simulations, we examine the tests' performance when the model is
misspecified. \readme{The plots have too many points; maybe we should only plot  every other diff value.  In the xyplot command, you could add the argument ``subset = diff \%\% .5 == 0''.}

@@

if(file.exists("sim1.rda")) {
  load("sim1.rda")
} else {
  sim1 <- simulation(sim="sim1", nobs=480, parms=c("loading","error","var"))
  sim1$nlevels <- factor(sim1$nlevels)
  levels(sim1$nlevels) <- paste("m=",levels(sim1$nlevels),sep="")

  sim1$nobs <- factor(sim1$nobs)
  levels(sim1$nobs) <- paste("n=",levels(sim1$nobs),sep="")
  levels(sim1$test) <- c("maxLM_O","WDM_O","LM_uO")
  save(sim1, file="sim1.rda")
}
sim1$test <- factor(as.character(sim1$test),
  levels = c("ordmax","ordwmax","catdiff"),
  labels = c("maxLM_o","WDM_o","LM_uo"))
  parlabs <- c(expression(lambda[11]), expression(phi[12]), 
               expression(psi[11]), expression(lambda[11] - lambda[62]), 
               expression(psi[11] - psi[66]))
  levels(sim1$pars) <- c("Loading1","Covariance","Error1","All Loadings","All Errors")  
@

\begin{figure}
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ test, the double-max test, and the unordered $\max\mathit{LM}$ test across three
      levels of the ordinal variable $m$ and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The parameter violating measurement invariance is $\lambda_{11}$.}
\label{fig:sim11res}
\setkeys{Gin}{width=\textwidth}
<<sim11res, fig=TRUE, height=7, width=7, echo=FALSE>>=
#sim1.tmp <- sim1[sim1$parms %in% c("loading")& sim1$pars %in% c("Loading1","All Loadings","All Errors"),]
#sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1,
             subset = (parms=="loading" & 
                      pars %in% c("Loading1","All Loadings","All Errors")),
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey,
             strip = function(..., which.given, factor.levels){
                 if(which.given==2){
                     strip.default(which.given, factor.levels = parlabs[c(1,4:5)], ...)
                 } else {
                     strip.default(which.given, factor.levels = factor.levels, ...)
                 }
             }))
@

@@

\begin{figure}
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the ordered double-max test the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is Covariance}
\label{fig:sim12res}
\setkeys{Gin}{width=\textwidth}
<<sim12res, fig=TRUE, height=7, width=7, echo=FALSE>>=
sim1.tmp <- sim1[sim1$test %in% c("maxLM_o","WDM_o","LM_uo")&sim1$parms %in% c("var") &sim1$pars %in% c("All Loadings", "Covariance","All Errors") ,]
sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

@@

\begin{figure}
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the ordered double-max test the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is Error1}
\label{fig:sim13res}
\setkeys{Gin}{width=\textwidth}
<<sim13res, fig=TRUE, height=7, width=7, echo=FALSE>>=
sim1.tmp <- sim1[sim1$test %in% c("maxLM_o","WDM_o","LM_uo")&sim1$parms %in% c("error") &sim1$pars %in% c("Error1","All Errors","All Loadings"),]
sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

@@

\seeme{I think it might be nicer to keep these plots all have three rows?(Not sure)}

@@

\subsection{Method}
\readme{Make sure there is space after each period and comma, the beginning of each sentence is capitalized, etc.}
Measurement invariance violations could occur in
each of the three parameters from Simulation 1 (factor loading, factor covariance, and unique variance), and they could also occur in the extra, unmodeled loading.  In each condition, a single parameter exhibited the violation.
Sample size and magnitude of measurement invariance violation were manipulated 
in the same way as they were in Simulation 1.The tested parameters were also the same as Simulation
1.

\subsection{Results}
Results are presented by the parameter that violated measurement invariance.
Of primary interest is Figure~\ref{fig:sim21res}, which displays results when the unmodeled loading violates 
measurement invariance. One can generally observe that 
tests of the first loading and error variance parameter exhibited high power, reflecting a high Type I error rate.  Tests associated with the factor covariance did not demonstrate this error, however.

\readme{Need some comparison between the unordered and ordered tests.}

When the modeled factor loading, $\lambda_{11}$, violated
measurement invariance, the statistics were generally able to pick up the violation despite the misspecification.
A similar pattern of specificity and power were observed when the unique variance and factor covariance paraameters violated measurement invariance; the figures associated with these results are shown in the Appendix.

\readme{We should say something about how the unordered test looks better when the unmodeled parameter violates MI.}
In summary, the proposed test statistics appear robust to unmodeled loading parameters,so long
as the umodeled parameter does not violate measurment invariance. If the unmodeled parameter 
dose violate measurement invariance, however, then the tests assign this violation to other 
parameters that do not violate measurement invariance. The impacted parameters include the error
variance and other loadings associated with the manifest variables that had the unmodeled 
loading. Thus, prior to testing measurement invariance, it is important to study the extent to
which the hypothesized model includes all parameters of importance.

@@

if(file.exists("sim2.rda")) {
  load("sim2.rda")
} else {
  sim2 <- simulation(sim="sim2", nobs=480, parms=c("extra","extra+loading","extra+var","extra+error"))

  save(sim2, file="sim2.rda")
}
sim2$test <- factor(as.character(sim2$test),
  levels = c("ordmax","ordwmax","catdiff"),
  labels = c("maxLM_o","WDM_o","LM_uo"))
  sim2$nlevels <- factor(sim2$nlevels)
  levels(sim2$nlevels) <- paste("m=",levels(sim2$nlevels),sep="")
  sim2$nobs <- factor(sim2$nobs)
  levels(sim2$nobs) <- paste("n=",levels(sim2$nobs),sep="")
  levels(sim2$pars) <- c("Loading1","Covariance","Error1","All Loadings","All Errors")

>>>>

Parameter invariance is a fundamental assumption for measurement model in 
developing psychometric scales. Violation of it could happen in any degree. 
Generally, researchers want to ignore minor violation parameters, e.g. random 
normally distributed deviations from a parameter value across groups. BSEM
(Bayesian Structural Equation Modelling) is proposed not to pick up this kind of 
violating parameters but only "real" violating ones by assigning priors instead of strictly 
constraining equality across groups. After identifying the violation parameters, 
those non-invariant parameters are freely estimated to ensure final correct estimation.

Instead of using this two-step Bayesian analysis procedure, we are applying a family of tests derived
from stochastic process. These tests have been recently proposed to identify specific parameters impacted by measurement invariance. However, they have only been used to a limited type of parameter violation. It is this paper's intent to provide a detailed examination 
of the tests' performance under practical research scenarios. Specifically, we will examine these tests
specificity in identifying the violation parameter without giving false positive results. Moreover, in practice model misspecification is an inevitable issue to a certain extent, we will demonstrate these tests
results under model misspecification condition. 
\seeme{What about this intro? too much BSEM?}
%In order to fix the psychometric scale, we need to specify the violating parameter.  
%Moreover, in practice model misspecification is an inevitable issue to a certain extent 
%and may interfere with parameter invariance specificity. A family of tests derived from 
%stochastic process has been recently proposed to identify specific parameters impacted 
%by measurement invariance violation. However, these tests have only been used to a limited 
%type of parameter violation. It is this paper's intent to provide a detailed examination 
%of the tests' performance under practical research scenarios.

@@

## remove cases with imputed values(not in 1,...,9)
compcases <- apply(YouthGratitude[,4:28], 1, function(x) all(x %in% 1:9))
yg <- YouthGratitude[compcases,]
@
Next, we fit two models in lavaan: a one-factor model where loadings are restricted 
to be equal across age groups, and a one-factor model where loadings are free across age groups.
<<modest>>=
## Restricted model
reduced <- cfa("f1 = ~gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
             data = yg, group = "agegroup", meanstructure = TRUE,
             group.equal = "loadings")

## Unrestricted model
full <- cfa("f1= ~gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
             data = yg, group = "agegroup", meanstructure = TRUE)
@
Finally, we can get the results of a likelihood ratio test via the \verb+anova()+ 
function, which implies that the GQ-6 violates measurement invariance.
<<LRT>>=
## LRT
anova(full, reduced)
@

Another similar statistics is $\mathit{LM}_\mathit{uo}$ as shown in Equation~ \eqref{eq:lmuo}. We can obtain this statistic by using the following function. 
<<LM_UO>>=
## unordered LM
lm_uo <- sctest(reduced, as.numeric(yg$agegroup), parm = 1:4, stat = "LMuo")
round(lm_uo$p.value, 4)
@
The above command specifies that we wish to test model parameter numbers 1 through 4, which are the factor loadings supplied by lavaan.  The ordering of parameters can be obtained via the command \verb+coef(reduced)+.

Because our sample size is large, this likelihood ratio test is known to be sensitive to 
small measurement invariance violations \cite{Ben80}.  That is, the above tests 
are sensitive to small measurement invariance violations that are not likely to be 
of interest to researchers.  For example, imagine that the 15-year-olds' 
parameters are slightly different than the other age groups.  
The 15-year-olds are in the middle of the age groups, and there 
is not likely to be any theoretical justification for 15-year-olds 
differing from every other age group.  One solution to this problem would be the {\em approximate invariance} methods proposed by Muth\'{e}n \cite{muthen13}\seeme{How to handle muthen's name in refs.bib file?}.
Alternatively, we can use the proposed family of score-based statistics to obtain tests that are sensitive to the ordering of age.

@@

<<Get Estimate>>=
dm <- sctest(reduced, as.numeric(yg$agegroup), parm = 1:4, stat = "DM")
cvm <- sctest(reduced, as.numeric(yg$agegroup), parm = 1:4, stat = "CvM")
maxlm <- sctest(reduced, as.numeric(yg$agegroup), parm = 1:4, stat = "maxLM")
round(c(dm$p.value, cvm$p.value, maxlm$p.value), 2)
@
%% move this code description to the cat2BB part. 
%In the code above, the \verb+parm=1:4+ argument means that 
%we wish to test model parameter numbers 1 through 4 (which are the factor loadings), 
%which are supplied by lavaan.  The ordering of parameters can be obtained via the 
%command \verb+coef(reduced)+.  
We see that two of the three $p$-values output at 
the end of the code are larger than that associated with the LRT (with the CvM statistic being non-significant).

@@

To carry out the tests, we can rely on the same function that we used
for the continuous test statistics.  As mentioned previously,
calculation of the ``maxLMo'' statistic (Equation~\eqref{eq:maxlmo})
can be lengthy from the need to simulate critical values.
%% This is more complicated than necessary, but it is written like
%% this so that we can load the ordinal critical values behind the
%% scenes without confusing the reader.
<<critvals,echo=FALSE,strip.white=all>>=
if(file.exists("opval1.rda")){
  load("opval1.rda")
} else {
  gefp1 <- gefp(reduced, fit = NULL, order.by = as.numeric(yg$agegroup),
                vcov = info_full, sandwich = FALSE, parm = 1:4)
  opval1 <- ordL2BB(gefp1)
  save(opval1, file="opval1.rda")
}
## Get test statistics using fnl=opval1 without showing it in the paper
wdmo <- sctest(reduced, as.numeric(yg$agegroup), parm = 1:4, stat = "WDMo")
maxlmo <- sctest(reduced, as.numeric(yg$agegroup), parm = 1:4, stat = "maxLMo",
                 fnl = opval1)
@ 
<<ordtest,eval=FALSE,strip.white=all>>=
wdmo <- sctest(reduced, as.numeric(yg$agegroup), parm = 1:4, stat = "WDMo")
maxlmo <- sctest(reduced, as.numeric(yg$agegroup), parm = 1:4, stat = "maxLMo")
@ 
<<pval,strip.white=all>>=
round(c(wdmo$p.value, maxlmo$p.value), 2)
@ 
In computing the ordinal test statistics, we
obtain~$p=\Sexpr{round(wdmo$p.value,2)}$
and~$p=\Sexpr{round(maxlmo$p.value,2)}$, respectively.
Both $p$-values are clearly larger than that of the 
likelihood ratio test and neither is significant at $\alpha=0.05$. 
This provides evidence that there is no measurement
invariance violation that is monotonic with age group.  Instead, given
the large sample size, the likelihood ratio test may be overly
sensitive to anomalous, non-monotonic violations at one (or a few) age groups.

@@

%Or alternatively, we could use the sctest.lavaan function, but assign ordinal statistics.
\begin{figure}
\caption{Fluctuation processes for the $\mathit{WDM}_o$ statistic (left panel) and the $\max\mathit{LM}_o$ statistic (right panel).}
\label{fig:ordres}
\setkeys{Gin}{width=\textwidth}
<<ordres, fig=TRUE, height=3, width=6, echo=FALSE, results=hide>>=
par(mfrow = c(1, 2), cex = 0.65)
sctest(reduced, as.numeric(yg$agegroup), parm = 1:4, stat = "WDMo", 
       fplot = TRUE, ylim = c(0,3.5), xlab = "Age group",
       ylab = "Weighted max statistics", axes = FALSE)
axis(1, at = 1:5, labels = c("10-11", "12-13", "14", "15", "16"))
box()

sctest(reduced, as.numeric(yg$agegroup), parm = 1:4, stat = "maxLMo", 
       fnl = opval1, fplot = TRUE, ylim = c(0,14), xlab = "Age group",
       ylab = "LM statistics", axes = FALSE)
axis(1, at = 1:5, labels = c("10-11", "12-13", "14", "15", "16"))
box()
@
\end{figure}

@@

To examine these specificity issues, we generated data from a
two-factor model with three indicators each (see Figure~\ref{fig:path}).
The measurement invariance
violation occurred in one of three places: the factor loading associated with Scale 1 
 ($\lambda_{11}$), the unique variance associated with Scale 1
 ($\epsilon_{11}$), or the factor covariance $\sigma_{12}$.  
We then tested for measurement invariance in seven subsets of parameters: the three
individual parameters noted above, all six factor loadings, % $\lambda_{11}--\lambda_{62}$, and 
and all six unique variances. % $\epsilon_{11}--\epsilon_{66}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Power and Type I error were examined across three sample sizes
($n=120,480,960$), three numbers of categories ($m=4,8,12$) and
17 magnitudes of invariance violations. 
The measurement invariance violations began at level $1+m/2$ of $V$
and were consistent thereafter. These
violations were based on the auxiliary variable $V$: individuals below
the $1+m/2$ percentile of $V$ deviated from individuals above the $1+m/2$ 
percentile  by $d$ times the parameters' asymptotic standard errors
(scaled by $\sqrt n$), with $d=0,0,25,0.5,...,4$. 
For each combination of sample size ($n$) $\times$ violation
magnitude ($d$) $\times$ violating parameter $\times$ categories $m$, 5000 datasets were
generated and tested. Statistics from Equations~\eqref{eq:wdm}, ~\eqref{eq:maxlmo} and~\eqref{eq:lmuo} were examined. 
In all conditions, we
maintained equal sample sizes at each level of the ordinal variable.

\subsection{Results}

Full simulation results are presented in Figure~\ref{fig:sim11res} to 
Figure~\ref{fig:sim13res}; 
Figure~\ref{fig:sim11res} displays power curves as a function of 
violation magnitude for the first
factor loading $\lambda_{11}$, with panels reflecting the parameters 
being tested and lines reflecting $n$.
Figures~\ref{fig:sim12res} and Figure~\ref{fig:sim13res} display 
similar power curves when the factor covariance 
$\phi_{12}$ and error variance $\epsilon_{11}$ violate measurement
invariance, respectively. 
All three test statistics exhibited 
similar results, with $\max\mathit{LM}_{uo}$ demonstrating lower power across all situations. This statistic 
is asymptotically equivalent to the LRT as what we describe in introduction. $\mathit{WDM}_o$  and $\max \mathit{LM}_o$ exhibited equivalent power curves because, when only one parameter is tested, the statistics are equivalent.

@@

In summary, we found that the proposed tests can attribute
measurement invariance violations to the correct parameter.  This
provides evidence that, in practice, one can have confidence in the
tests' abilities to locate the measurement invariance violation.  Of
course, this statement is qualified by the fact that, in this
simulation, the model was correctly specified.  In the following
simulations, we examine the tests' performance when the model is
misspecified.

@@

if(file.exists("sim1.rda")) {
  load("sim1.rda")
} else {
  sim1 <- simulation(sim = "sim1", nobs = 480, parms = c("loading", "error", "var"))
  sim1$nlevels <- factor(sim1$nlevels)
  levels(sim1$nlevels) <- paste("m=", levels(sim1$nlevels), sep="")

  sim1$nobs <- factor(sim1$nobs)
  levels(sim1$nobs) <- paste("n=", levels(sim1$nobs), sep="")
  levels(sim1$test) <- c("maxLM_O", "WDM_O", "LM_uO")
  save(sim1, file="sim1.rda")
}
sim1$test <- factor(as.character(sim1$test),
  levels = c("ordmax", "ordwmax", "catdiff"),
  labels = c("maxLM_o", "WDM_o", "LM_uo"))
  parlabs <- c(expression(lambda[11]), expression(phi[12]), 
               expression(psi[11]), expression(lambda[11] - lambda[62]), 
               expression(psi[11] - psi[66]))
  levels(sim1$pars) <- c("Loading1", "Covariance", "Error1", "All Loadings", "All Errors")  
@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and $\mathit{LM}_{uo}$ test across three levels of the ordinal variable $m$ and measurement invariance 
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The parameter violating measurement 
invariance is $\lambda_{11}$.}
\label{fig:sim11res}
\setkeys{Gin}{width=\textwidth}
<<sim11res, fig=TRUE, height=7, width=7, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
             subset = (parms == "loading" & 
                       pars %in% c("Loading1", "All Loadings", "All Errors") & diff %% 0.5 == 0),
             type = "b", xlab = "Violation Magnitude", ylab = "Power", key=mykey,
             strip = function(..., which.given, factor.levels){
                 if(which.given == 2){
                     strip.default(which.given, factor.levels = parlabs[c(1, 4:5)], ...)
                 } else {
                     strip.default(which.given, factor.levels = factor.levels, ...)
                 }
             }))
@

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ tests, $\mathit{WDM}_o$ test $\mathit{LM}_{uo}$ across three levels of the ordinal variable $m$, and measurement invariance violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is $\sigma_{12}$}
\label{fig:sim12res}
\setkeys{Gin}{width=\textwidth}
<<sim12res, fig=TRUE, height=7, width=7, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
             subset = (parms == "var" & 
                       pars %in% c("Covariance", "All Loadings", "All Errors") & diff %% 0.5 == 0),
             type = "b", xlab = "Violation Magnitude", ylab="Power", key = mykey,
             strip = function(..., which.given, factor.levels){
                 if(which.given == 2){
                     strip.default(which.given, factor.levels = parlabs[c(2, 4:5)], ...)
                 } else {
                     strip.default(which.given, factor.levels = factor.levels, ...)
                 }
             }))
@

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and $\mathit{LM}_{uo}$ test across three levels of the ordinal variable $m$ and measurement invariance 
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is $\psi{11}$}
\label{fig:sim13res}
\setkeys{Gin}{width=\textwidth}
<<sim13res, fig=TRUE, height=7, width=7, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
             subset = (parms == "error" & 
                       pars %in% c("Error1", "All Loadings", "All Errors") & diff %% 0.5 == 0),
             type = "b", xlab = "Violation Magnitude", ylab = "Power", key = mykey,
             strip = function(..., which.given, factor.levels){
                 if(which.given == 2){
                     strip.default(which.given, factor.levels = parlabs[c(3,4:5)], ...)
                 } else {
                     strip.default(which.given, factor.levels = factor.levels, ...)
                 }
             }))
@

@@

@@

\subsection{Method}
Measurement invariance violations could occur in
each of the three parameters from Simulation 1 (factor loading, factor covariance, and unique variance), and they could also occur in the extra, unmodeled loading. In each condition, a single parameter exhibited the violation.
Sample size and magnitude of measurement invariance violation were manipulated 
in the same way as they were in Simulation 1. The tested parameters were also the same as Simulation
1.

\subsection{Results}
Results are presented by the parameter that violated measurement invariance.
Of primary interest is Figure~\ref{fig:sim21res}, which displays results when the unmodeled loading violates 
measurement invariance. One can generally observe that 
tests of the first loading and error variance parameter exhibited high power, reflecting a high Type I error rate.  Tests associated with the factor covariance did not demonstrate this error, however. In terms of specific statistic performance, $\max \mathit{LM}_o$ test and $\mathit{WDM}_o$ test demonstrated higher power than $\mathit{LM}_{uo}$ in each panel, especially with increasing levels. It is indicating ordered statistics is more inclined to pick up the unmodelled parameter violation in the manifest variables' associated loading and error terms. 
\seeme{Is this sentence enough to compare ordered and unordered statistics?}

When the modeled factor loading, $\lambda_{11}$, violated
measurement invariance, the statistics were generally able to pick up the violation despite the misspecification.
A similar pattern of specificity and power were observed when the unique variance and factor covariance parameters violated measurement invariance; the figures associated with these results are shown in the Appendix. Also, the power of ordered statistics $\max \mathit{LM}_o$ test and $\mathit{WDM}_o$ test showed higher power than unordered $\mathit{LM}_{uo}$ in each panel.

In summary, the proposed test statistics appear robust to unmodeled loading parameters,so long
as the umodeled parameter does not violate measurement invariance. If the unmodeled parameter 
dose violate measurement invariance, however, then the tests assign this violation to other 
parameters that do not violate measurement invariance. The impacted parameters include the error
variance and other loadings associated with the manifest variables that had the unmodeled 
loading. Thus, prior to testing measurement invariance, it is important to study the extent to
which the hypothesized model includes all parameters of importance.

@@

if(file.exists("sim2.rda")) {
  load("sim2.rda")
} else {
  sim2 <- simulation(sim = "sim2", nobs = 480, 
  parms = c("extra", "extra+loading", "extra+var", "extra+error"))

  save(sim2, file="sim2.rda")
}
sim2$test <- factor(as.character(sim2$test),
  levels = c("ordmax", "ordwmax", "catdiff"),
  labels = c("maxLM_o", "WDM_o", "LM_uo"))
  sim2$nlevels <- factor(sim2$nlevels)
  levels(sim2$nlevels) <- paste("m=", levels(sim2$nlevels), sep="")
  sim2$nobs <- factor(sim2$nobs)
  levels(sim2$nobs) <- paste("n=", levels(sim2$nobs), sep="")
  parlabs <- c(expression(lambda[11]), expression(phi[12]), 
               expression(psi[11]), expression(lambda[11] - lambda[62]), 
               expression(psi[11] - psi[66]))
  levels(sim2$pars) <- c("Loading1", "Covariance", "Error1", "All Loadings", "All Errors")
